---
title: "spotify_cleaning"
format: html
---

```{r}
# load libraries
library(jsonlite)
library(tidyverse)
```

```{r}
# read in data as json
spotify_json <- read_json("../data/raw/StreamingHistory_music_0.json", simplifyVector = TRUE)

# check for any nested lists that weren't converted to vectors
spotify_json |> 
  select(where(is.list)) |> 
  glimpse() # i think we're good to go?

# write to csv to for future use
# don't think we have to worry about flattening since there don't seem to be any unconverted lists
write_csv(spotify_json, "../data/processed/spotify.csv")
```

```{r}
# read in csv
spotify <- read_csv("../data/processed/spotify.csv")

# convert entire dataset to Mountain Standard Time
spotify <- spotify |> 
  mutate(endTime = with_tz(endTime, tz = "UTC")) |>  # should be UTC to begin with but ensure that it is
  mutate(endTime = with_tz(endTime, tz = "America/Phoenix"))
```

```{r}
# split endTime into separate variables; rename and reorder other variables
spotify <- spotify |> 
  mutate(date = date(endTime),
         end_time = hms::as_hms(endTime)) |> 
  rename(artist = artistName,
         track = trackName,
         ms_played = msPlayed) |> 
  relocate(c(date, end_time, track), .before = artist)
```

```{r}
# subset only songs during June AZ visit
spotify_jun <- spotify |> 
  filter(endTime >= ymd_hms("2026-06-12 17:57:00", tz = "America/Phoenix") 
         & endTime <= ymd_hms("2026-06-15 08:56:00", tz = "America/Phoenix"))


# subset only songs during July AZ visit
# times seem to be off by an hour (maybe for just some? - hard to tell because time on phone photos is sometimes inaccurate too, so nothing to compare to)
# it is what it is
spotify_july <- spotify |> 
  filter(endTime >= ymd_hms("2026-07-17 17:27:00", tz = "America/Phoenix") 
         & endTime <= ymd_hms("2026-07-22 07:37:00", tz = "America/Phoenix"))

spotify_az <- spotify |> 
  filter(endTime >= ymd_hms("2026-06-12 17:57:00", tz = "America/Phoenix") &
           endTime <= ymd_hms("2026-06-15 08:56:00", tz = "America/Phoenix") | 
           endTime >= ymd_hms("2026-07-17 17:27:00", tz = "America/Phoenix") & 
           endTime <= ymd_hms("2026-07-22 07:37:00", tz = "America/Phoenix"))

head(spotify_jun)
head(spotify_july)
head(spotify_az)
```




## Resources

- https://datacarpentry.github.io/r-socialsci/07-json.html
- https://github.com/rstudio/cheatsheets/blob/main/lubridate.pdf
- https://dplyr.tidyverse.org/reference/between.html
- https://en.wikipedia.org/wiki/Time_in_Arizona
- "I am working with Spotify data. I want to the filter the dataset to a range of datetimes based on a variable called endTime, which is of S3: POSIXct class. I only want dates after 2026-06-12 17:57:00. Here is my code (spotify is a tibble that displays as expected): spotify |> filter(endTime >= "2026-06-12 17:57:00") The data is successfully filtered to the date of 2026-06-12, but it gives observations before 17:57:00 on that day. What am I doing wrong? Ask clarifying questions if necessary, and do not be sycophantic." 
